Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-02-05 paper/paper.tex
fixed eqref in appendix and incorporated coauthors edits

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that the key merit of $\cplx$-valued neural
networks is simultaneous phase rotation and amplitude adjustment, intrinsic to multiplication
in $\cplx$ field. Despite inherently higher arithmetic complexity, which stems from up to four
floating point operations per single $\cplx$ multiplication, \citep{monning_evaluation_2018},
$\cplx$-arithmetic reduces ineffective degrees of freedom in comparison with an analogous
double-dimensional $\real$-valued network. \citet{hirose_complex-valued_2009} list numerous
applications related to acoustic and radio signal processing and demonstrate superiority of
$\cplx$-networks in the task of shallow landmine detection using ground penetrating radar
imaging. More recently they were applied to music transcription and speech recognition
\citep{trabelsi_deep_2017}. \citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs
to deal with exploding~/~vanishing gradients and apply the recurrent models to long-term
dependence prediction tasks. \citet{wisdom_full-capacity_2016} expand on their work by optimizing
recurrent weights on the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018}
investigate different $\cplx$-valued gating mechanisms unitary-RNNs and conduct experiments
on human motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification.

@@

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
have shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated multitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Bayesian inference is a very rich field of research with a large body of research and
techniques that enable practical quantification of predictive, parameter, or model
uncertainty and its propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Under suitable modelling and prior assumptions, Bayesian methods
can be used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$, \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that despite dropout showing more variable
performance, than magnitude pruning, it nevertheless induces non-uniform sparsity throughout
the model, which is essential for superior compression \citep{he_amc:_2018}.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational sparsification and stochastic sparsification techniques we
extend Bayesian dropout for $\cplx$-valued networks. We assess the performance-compression
trade-off of the the proposed Bayesian technique by conducting a large-scale numerical
study of $\cplx$-valued networks on two tasks: image recognition on MNIST-like and CIFAR10
datasets and music transcription on MusicNet, \citep{thickstun_learning_2017}. We reproduce
the state-of-the-art result by \citet{trabelsi_deep_2017} for $\cplx$-valued networks on
the latter with $\approx 1-2\%$ of remaining non-zero parameters.

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate and compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes.

@@

The posterior itself is intractable, save for the relatively simple cases \citep{citation_needed},
and therefore exact inference is traded for tractability and scalability offered by Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
%
The natural choice for $\rho$ is Kullback-Leibler divergence of density $q$ from $p$,
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
the main reason being that it satisfies the following identity for well-behaved densities:
% $(q, p, \pi)$
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which
makes it possible to employ stochastic gradient-based optimization methods to solve
\eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
% (DSVI) titsias-lazaro.tex
\todo{DSVI}
\citep{titsias_doubly_2014}

@@

With a special family of posterior approximation, variational inference can be used as
a regularization method and as model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect \citep{wan_regularization_2013},
and Gaussian dropout \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. The divergence term in \eqref{eq:elbo} unravels into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for truncated $\alpha \in (0, 1)$. \citet{molchanov_variational_2017}
refine the approximation by using a nonlinear regression based on sigmoid and soft-plus.
% see p.6 sec 3.4 of kingma_variational_2015
%
In sec.~\ref{sub:real-chisq-grad} we verify the approximation against Monte-Carlo
estimates of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic
grid, and confirm its accuracy, by comparing its derivative to the exact derivative of
\eqref{eq:improper-kl-div-real} with respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another variant of Bayesian
dropout, called \textit{Automatic Relevance Determination}, is developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\citep{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

>>>>

Inherently complex-valued nature of data in signal processing, specifically, acoustic or radio
frequency signal analysis, has been the key driver behind adoption of $\cplx$-networks.
%
\citet{hirose_complex-valued_2009} argues that the key merit of $\cplx$-valued neural
networks is simultaneous phase rotation and amplitude adjustment, intrinsic to multiplication
in $\cplx$ field. Despite inherently higher arithmetic complexity, which stems from up to four
floating point operations per single $\cplx$ multiplication \citep{monning_evaluation_2018}
$\cplx$-arithmetic reduces ineffective degrees of freedom in comparison with an analogous
double-dimensional $\real$-valued network. \citet{hirose_complex-valued_2009} list numerous
applications related to acoustic and radio signal processing and demonstrate superiority of
$\cplx$-networks in the task of shallow landmine detection using ground penetrating radar
imaging. More recently they were applied to music transcription and speech recognition
\citep{trabelsi_deep_2017}. \citet{arjovsky_unitary_2016} propose $\cplx$-valued unitary-RNNs
to deal with exploding~/~vanishing gradients and apply the recurrent models to long-term
dependence prediction tasks. \citet{wisdom_full-capacity_2016} expand on their work by optimizing
recurrent weights on the $\cplx^{n \times n}$ Stiefel manifold. And \citet{wolter_complex_2018}
investigate different $\cplx$-valued gating mechanisms unitary-RNNs and conduct experiments
on human motion prediction and music transcription. \citet{yang_complex_2019} develop complex
transformer, with $\cplx$-attention and complex encoder-decoder, and apply it to music
transcription and wireless signal classification.

@@

% on dropout
Dropout \citep{hinton_improving_2012} prevents overfitting by injecting multiplicative
binary noise into layer's weights, which breaks up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
have shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated multitask learning, then by sampling uncorrelated weight masks dropout acts
as a superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% segue into bayesian inference
Bayesian inference is a very rich field of research with a large body of research and
techniques that enable practical quantification of predictive, parameter, or model
uncertainty and its propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Bayesian methods with suitable modelling and prior assumptions, such
as Variational Dropout, can be used towards model sparsification
\citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$ \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that, despite dropout showing more variable
performance than magnitude pruning, it induces non-uniform sparsity throughout the model,
which has been shown by \citet{he_amc:_2018} to be essential for superior compression.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational sparsification and stochastic sparsification techniques we
extend Bayesian dropout for $\cplx$-valued networks. We assess the performance-compression
trade-off of the proposed Bayesian technique by conducting a large-scale numerical study
of $\cplx$-valued networks on two tasks: image recognition on MNIST-like and CIFAR10 datasets
and music transcription on MusicNet \citep{thickstun_learning_2017}. We reproduce the
state-of-the-art result by \citet{trabelsi_deep_2017} for $\cplx$-valued networks on the
latter with $\approx 5\%$ of remaining non-zero parameters.

The paper is structured as follows. In sec.~\ref{sec:bayesian_dropout} we review Bayesian
dropout techniques, and in sec.~\ref{sec:c_valued_networks} we provide a brief summary of
the inner working of complex-valued networks as functions of complex argument. The main
contribution of this study is presented in sec.~\ref{sec:dropout_for_c_valued_layers},
where we consider different variational approximations and priors, outline the tricks
and derive expressions for the divergence term. In sec.~\ref{sec:experiments} we evaluate
the sparsification rate, compare the resulting performance of various $\cplx$-networks
proposed in prior work, and discuss the outcomes.

@@

The posterior distribution itself is intractable, save for the relatively simple cases, and
therefore exact inference is traded for tractability and scalability offered by Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$ \citep{williams_simple_1992} or can be represented as differentiable transformations
of non-parametric noise \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
%
The natural choice for $\rho$ is Kullback-Leibler divergence of density $q$ from $p$,
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
the main reason being that it satisfies the following identity for well-behaved densities:
% $(q, p, \pi)$
\begin{multline}  \label{eq:kl-div-master}
  % \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
  \log p(D)
    - KL(q \| p(\omega \mid D))
    \\ = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q \| \pi)
    \,.
\end{multline}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \mathbb{E}_{\omega \sim q_{\theta}}
        \log p_{\phi}(D \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$ \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citet{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which
makes it possible to employ stochastic gradient-based optimization methods to solve
\eqref{eq:variational-progam}. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
% (DSVI) titsias-lazaro.tex \todo{DSVI} \citep{titsias_doubly_2014}

@@

With a special family of posterior approximation, variational inference can be used as
a regularization method and as a model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect \citep{wan_regularization_2013},
and Gaussian dropout \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. The divergence term in \eqref{eq:elbo} unravels into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta \| \pi)
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for truncated $\alpha \in (0, 1)$. \citet{molchanov_variational_2017}
refine the approximation by using a nonlinear regression based on sigmoid and soft-plus.
% see p.6 sec 3.4 of kingma_variational_2015
We independently verify their approximation against Monte-Carlo estimates for $\alpha$
varying over a fine logarithmic grid and the exact expression for the derivative of
\eqref{eq:improper-kl-div-real}, which we derive in appendix~\ref{sub:real-chisq-grad}.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another variant of Bayesian
dropout, called \textit{Automatic Relevance Determination}, is developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes},
fits a prior distribution while performing variational inference. Maximizing \eqref{eq:elbo}
over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \citep{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \citep{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\citet{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \citet{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}. They argue for rejecting the $(\mu, \alpha)$
parameterization in favour of the equivalent $(\mu, \sigma^2)$, since the latter renders $
  \nabla_\mu \tilde{\mathcal{L}}
$ independent from the local output noise, injected by the trick \eqref{eq:r-gauss-trick},
and further stabilizes gradients in SGVB \eqref{eq:elbo}. The relevance $\alpha$ is
calculated using the ratio $
  \tfrac{\sigma^2}{\mu^2}
$ when needed in the divergence term.
